---
title: "Simulation Test Thesis"
output: html_document
date: "2023-10-27"
---

---
title: "Simulation Test"
output: html_document
date: "2023-10-25"
---
```{r setup}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
## Setup
### Libraries
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(caTools)
library(plotly)
library(MASS)
library(matlib)
library(dplyr)
library(pROC)
library(parallel)
library(CalibrationCurves)
source("R/Functions.R")
```
### Initialize variables
```{r}
# Set random seed for reproducibility
set.seed(28)
p      <- 0                             # proportion of correlated predictors
npred  <- 3                             # number of predictors
z      <- 0.2                           # z 
n      <- 1000000                       # Number of samples in the dataset
a <- 4                                  # the effect size
```

## Simulate data
Simulate data for diseased class and non-diseased class using a covariance matrix and mean vector using the MASS package. The covariance matrix is chosen such that the two predictors are correlated. The mean vector is chosen such that the diseased class has a higher mean than the non-diseased class for both predictors. The data is then combined into a single dataset.
```{r, cache=T}
# mean structures
mu0 <- c(rep(1,npred))       # vector: class 0 means

# covariance structures
sigma0 <- scenario()         # class 0
```

Experiment 1
```{r, cache=T}
# Generate data
data_experiment3 <- mvrnorm(n, mu = mu0, Sigma = sigma0) %>%
  as.data.frame()

# Rename columns to "predictor1", "predictor2" ,...
colnames(data_experiment3) <- c(paste0("predictor", 1:(ncol(data_experiment3))))

# Simulate binary y to have Prob(y=1) = 1/[1+exp(-L)]
L1 <-  0.2*a*data_experiment3$predictor1 + 0.2*a*data_experiment3$predictor2 + 0.1*a*data_experiment3$predictor3
mean(L1)
L1 <- - mean(L1) + 0.2*a*data_experiment3$predictor1 + 0.2*a*data_experiment3$predictor2 + 0.1*a*data_experiment3$predictor3 # 50/50 split

y <- ifelse(runif(n) < plogis(L1), 1, 0)
data_experiment3 <- cbind(data_experiment3, y)

#proportion y's in data
prop.table(table(data_experiment3$y))
```

Experiment 2
```{r, cache=T}
# Generate data
data_experiment4 <- mvrnorm(n, mu = mu0, Sigma = sigma0) %>%
  as.data.frame()

# Rename columns to "predictor1", "predictor2" ,...
colnames(data_experiment4) <- c(paste0("predictor", 1:(ncol(data_experiment4))))

# Simulate binary y to have Prob(y=1) = 1/[1+exp(-L)] 0.2255
L2 <-  0.2*a*data_experiment4$predictor1 + 0.2*a*data_experiment4$predictor2
mean(L2)
L2 <- - mean(L2) + 0.2*a*data_experiment4$predictor1 + 0.2*a*data_experiment4$predictor2 # 50/50 split

y <- ifelse(runif(n) < plogis(L2), 1, 0)
data_experiment4 <- cbind(data_experiment4, y)

#proportion y's in data
prop.table(table(data_experiment4$y))
```
# Save Data in RDS Format in data folder
```{r}
saveRDS(data_experiment3, "data/simulation/data experiment 3.rds")
saveRDS(data_experiment4, "data/simulation/data experiment 4.rds")
```


## Plot distributions for Predictors
### 2D Histograms
```{r, cache=TRUE}
# Plot for Predictor 1
ggplot(data_experiment3, aes(x = predictor1, fill = as.factor(y))) +
  geom_histogram(alpha = 0.5, position = 'identity', bins = 200) +
  ggtitle("Distribution of Diseased and Non-Diseased Classes for Predictor 1")

# Plot for Predictor 2
ggplot(data_experiment3, aes(x = predictor2, fill = as.factor(y))) +
  geom_histogram(alpha = 0.5, position = 'identity', bins = 200) +
  ggtitle("Distribution of Diseased and Non-Diseased Classes for Predictor 2")

# Plot for Predictor 3
ggplot(data_experiment3, aes(x = predictor3, fill = as.factor(y))) +
  geom_histogram(alpha = 0.5, position = 'identity', bins = 200) +
  ggtitle("Distribution of Diseased and Non-Diseased Classes for Predictor 3")
```

### 2D Histograms
```{r, cache=TRUE}
# Plot for Predictor 1
ggplot(data_experiment4, aes(x = predictor1, fill = as.factor(y))) +
  geom_histogram(alpha = 0.5, position = 'identity', bins = 200) +
  ggtitle("Distribution of Diseased and Non-Diseased Classes for Predictor 1")

# Plot for Predictor 2
ggplot(data_experiment4, aes(x = predictor2, fill = as.factor(y))) +
  geom_histogram(alpha = 0.5, position = 'identity', bins = 200) +
  ggtitle("Distribution of Diseased and Non-Diseased Classes for Predictor 2")

# Plot for Predictor 3
ggplot(data_experiment4, aes(x = predictor3, fill = as.factor(y))) +
  geom_histogram(alpha = 0.5, position = 'identity', bins = 200) +
  ggtitle("Distribution of Diseased and Non-Diseased Classes for Predictor 3")
```

### Test models
Model 3
```{r, cache=TRUE}
model3 <- glm(y ~ ., data = data_experiment3, family = binomial)
summary(model3)
predictions3 <- predict(model3, newdata = data_experiment3, type = "response")

# Assuming 'colAUC' is available or replaced with an equivalent function
colAUC(predictions3, data_experiment3$y, plotROC = F)
```
Model 4
```{r, cache=TRUE}
model4 <- glm(y ~ ., data = data_experiment4, family = binomial)
summary(model4)
predictions4 <- predict(model4, newdata = data_experiment4, type = "response")

# Assuming 'colAUC' is available or replaced with an equivalent function
colAUC(predictions4, data_experiment4$y, plotROC = F)
```