---
title: "Simulation_Test2"
output: html_document
date: "2023-10-27"
---

---
title: "Simulation Test"
output: html_document
date: "2023-10-25"
---

## Libraries
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(caTools)
library(plotly)
library(MASS)
library(matlib)
```

## Initialize variables
```{r}
# Set random seed for reproducibility
set.seed(123)
p      <- 1                               # proportion of correlated predictors
npred  <- 3                               # number of predictors
z      <- 0.2                             # z 
n      <- 100000                            # Number of samples in each class
a <- 2                                    # the effect size
```

calculate sigma0 and sigma1
```{r}
scenario <- function(){
  # set up correlations
  corr0  <-  matrix(0, npred, npred)         # matrix: set up for cov matrix, 0 on diagonals
  corr0[1:npred*p, 1:npred*p] = z            # class 0 
  diag(corr0) = 0
  # covariance structures
  sigma0 <-  diag(npred)  + corr0            # matrix: cov matrix of class 0 
  return(sigma0)
}
```


## Simulate data
Simulate data for diseased class and non-diseased class using a covariance matrix and mean vector using the MASS package. The covariance matrix is chosen such that the two predictors are correlated. The mean vector is chosen such that the diseased class has a higher mean than the non-diseased class for both predictors. The data is then combined into a single dataset.
```{r}
# mean structures
mu0 <- c(rep(1,npred))       # vector: class 0 means

# covariance structures
sigma0 <- scenario()         # class 0

# Generate data for non-diseased class
data <- mvrnorm(n, mu = mu0, Sigma = sigma0) %>%
  as.data.frame()

# Rename columns to "predictor1", "predictor2" ,...
colnames(data) <- c(paste0("predictor", 1:(ncol(data))))

# Simulate binary y to have Prob(y=1) = 1/[1+exp(-L)]
L <-  0.4*a*data$predictor1 + 0.2*a*data$predictor2 + 0.1*a*data$predictor3 + 0.1*a*data$predictor1*data$predictor2

L <- - mean(L) + 0.4*a*data$predictor1 + 0.2*a*data$predictor2 + 0.1*a*data$predictor3 + 0.1*a*data$predictor1*data$predictor2 # 50/50 split

y <- ifelse(runif(n) < plogis(L), 1, 0)
data <- cbind(data, y)

#proportion y's in data
prop.table(table(data$y))
```

## Plot distributions for Predictor 1 and Predictor 2
```{r}
# Plot for Predictor 1
ggplot(data, aes(x = predictor1, fill = as.factor(y))) +
  geom_histogram(alpha = 0.5, position = 'identity', bins = 30) +
  ggtitle("Distribution of Diseased and Non-Diseased Classes for Predictor 1")

# Plot for Predictor 2
ggplot(data, aes(x = predictor2, fill = as.factor(y))) +
  geom_histogram(alpha = 0.5, position = 'identity', bins = 30) +
  ggtitle("Distribution of Diseased and Non-Diseased Classes for Predictor 2")
```


```{r, warning=FALSE}
# Calculate densities for class 1 and class 0
density1 <- with(data[data$y == 1, ], kde2d(predictor1, predictor2, n = 30))
density0 <- with(data[data$y == 0, ], kde2d(predictor1, predictor2, n = 30))

# Create the interactive 3D plot
p1 <- plot_ly(z = ~density1$z) %>% 
  add_surface(
    x = ~density1$x,
    y = ~density1$y,
    colorscale = list(c(0, "red"), list(1, "pink")),
    opacity = 0.8
  )

p2 <- plot_ly(z = ~density0$z) %>% 
  add_surface(
    x = ~density0$x,
    y = ~density0$y,
    colorscale = list(c(0, "blue"), list(1, "lightblue")),
    opacity = 0.8
  )

plot <- subplot(p1, p2, nrows = 1, shareX = TRUE, shareY = TRUE)
plot
```


## Logistic Regression and checking AUC (to see if it is correct)
```{r, warning=F}
n_iter <- 800                              # Number of iterations for logistic regression
auc_values <- numeric(n_iter)              # Vector to store AUC values
sample_ratio <- 0.2                        # Proportion of data to use in each logistic regression

# Run logistic regression and calculate AUC for n_iter iterations
for(i in 1:n_iter) {
  # Sample a subset of the data
  sample_index <- sample(1:nrow(data), sample_ratio * nrow(data))
  sample_data <- data[sample_index, ]
  
  # Run logistic regression with two predictors
  model <- glm(y ~ . + predictor1*predictor2, data = sample_data, family = binomial)
  
  # Predict on the same sample set
  predictions <- predict(model, newdata = sample_data, type = "response")
  
  # Calculate AUC
  auc <- colAUC(predictions, sample_data$y, plotROC = FALSE)
  auc_values[i] <- auc  # Directly store the AUC value
  
  # Save the model predictors and coefficients to calculate average coefficients
  if(i == 1) {
    coefficients <- model$coefficients
  } else {
    coefficients <- coefficients + model$coefficients
  }
  
}

# Calculate average AUC
average_auc <- mean(auc_values)
print(paste("Average AUC: ", round(average_auc, 2)))

# Calculate average coefficients
average_coefficients <- coefficients / n_iter
print(paste("Average coefficients: ", round(average_coefficients, 2)))
```
